```{r}
# load libraries
library(tidyverse)
library(readxl)
library(janitor)

smoke_complete <- read_excel("data/smoke_complete.xlsx",
                             sheet=1,
                             na="NA")
```

# mutate()
```{r}
smoke_complete %>%
    mutate(age_at_death = (age_at_diagnosis + days_to_death)/365)
```
## combining functions --- signif()
```{r}
smoke_complete %>%
    mutate(cigarettes_total =  cigarettes_per_day* days_to_death) %>%
    mutate(cigarettes_total = signif(cigarettes_total, 1))
```
###  What is a factor??? part 1.

# tabyl()
- from the janitor package
```{r}
smoke_complete %>%
    #reassign the gender variable to be a factor
    mutate(gender = factor(gender, levels = c("female", "male"))) %>%
    janitor::tabyl(gender) # we do this to point to the janitor package. In case we had another package with the same name
```
###quick cross table
```{r}
smoke_complete %>%
    tabyl(disease, gender)

```
# mutate() "case_when"
*condition* ~ *category_name*
```{r}
smoke_complete %>%
    mutate(
        cigarettes_category = case_when(
          cigarettes_per_day <= 5 ~ '0-5',
          cigarettes_per_day >5 & cigarettes_per_day <=10 ~ '6-10',
          cigarettes_per_day > 10 ~ '11+'
            )
        ) %>%
    mutate(cigarettes_category =
               factor(cigarettes_category,
                      levels = c("0-5", "6+")
                      )
           ) %>%
    janitor::tabyl(cigarettes_category)
```
### mutate_if() - look into this

# summarize()
- produces summaries across the data
Center: `mean()`, `median()`
Spread: `sd()` (standard deviation)
Range: `min()`, `max()`
Position: `first()`, `last()`, `nth()`
Count: `n()`, `n_distinct()`

```{r}
smoke_complete %>%
    summarize(average_cigarettes = mean(cigarettes_per_day))

```

#group_by()/summarize()
- splits data into separate dataframes by a category
- Usually followed by summarize()
```{r}
smoke_complete %>%
   group_by(gender, disease) %>%
   summarize(mean_cigarettes = mean(cigarettes_per_day),
            sd = sd(cigarettes_per_day))
```
# count()
- use group_by to count things
```{r}
smoke_complete %>%
    group_by(disease) %>%
    summarize(count = n())
```
