# Load required libraries
```{r}
library(tidyverse)
library(readxl)
```
##  read_excel()
```{r}
brca_clinical <- read_excel("data/fileshow.xlsx",
                            sheet = 1,
                            skip = 1,
                            na = "NA"
                            )
```
##  dim()
##  head()
## colnames()
##  View()
## summary()
## glimpse()
The output provided by `glimpse()` includes:

- data structure: `data.frame`
- dimensions:
- column-by-column information: each prefaced with a `$`, and includes the column name, data type (num, int, Factor)

# ggplot2

A `ggplot2` graphic consists of a:

- `mapping` of variables in `data` to
- `aes()`thetic attributes of
- `geom_`etric objects.

```{r}
ggplot(data = brca_clinical) +
  aes(
    x = days_to_last_follow_up / 365,
    y = days_to_birth
    color = gender
    ) +
  geom_point() +

  ggtitle("Age versus Days to Last Follow Up") +

  xlab("Age at Diagnosis (Years)") +
  ylab("Days to Follow Up")

```
## Save
```{r}
ggsave("follow_up_plot.jpg")
```
