```{r}
library(tidyverse)
library(readxl)

smoke_complete <- readexcel("data/smoke_complete.xlsx",
sheet = 1,
na = "NA")

# remove some columns - we'll talk about this later
smoke_complete <- smoke_complete %>%
  select(age_at_diagnosis, tumor_stage,
         cigarettes_per_day, gender,
         vital_status, disease)
```
## Customizing a Scatterplot
```{r}
our_plot <- ggplot(smoke_complete) +

  aes(x = age_at_diagnosis,
      y = cigarettes_per_day,
      color = disease) +

  geom_point() +

  labs(title = "Cigarettes per Day versus Age at Diagnosis",
       x = "Age at Diagnosis",
       y = "Cigarettes Smoked per Day")

our_plot
```
# Changing visual properties using built in themes
```{r}
our_plot +
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45))
```
## Themes: [here](https://ggplot2.tidyverse.org/reference/ggtheme.html)


# Categorical Data
- Factors: easier to organize categories. We will touch on this in the next chapter

# Boxplots

```{r}
ggplot(smoke_complete) +

  aes(x = vital_status,
      y = cigarettes_per_day,
      fill = vital_status) +

  geom_boxplot()
```
## Adding subplots
- Facet wrap requires "disease" to be wrapped in vars()
- We also added a limit on the y axis
- Puts data on one scale which is nice

```{r}
ggplot(smoke_complete) +

  aes(x = vital_status,
      y = cigarettes_per_day,
      fill = vital_status) +

  geom_boxplot() +

  ylim(c(0,20)) +

  facet_wrap(vars(disease))
```

# Data Manipulation with dplyr

- `arrange()` - sorting a `data.frame` by a variable
  - Descending order: arrange(desc()). Or just put "-" in front of the name
  - Can also sort by multiple variables
- `filter()` - subsetting a `data.frame` by criteria
  - ex: filter(cigarettes_per_day < 20)
  Chain using "&" (AND) or "|" (OR)
- `select()` - returning only a few columns from our `data.frame`
  - "-" to select everything but the variable
- `group_by()/summarize()` - summarizing our data.frame, such as counting and computing means
- `mutate()` - transforming variables in our data
- `%>%` - the pipe character, which lets us join our verbs together in a *pipeline*.

## Examples
```{r}
smoke_complete %>%
  filter(gender == "male" |
           tumor_stage == "stage iv")

# Multiple categories "%in%" operator
smoke_complete %>%
  filter(tumor_stage %in% c("stage ia", "stage ib"))

# Everything but gender
smoke_complete %>%
  select(-gender)
```
## The difference between `filter()` and `select()`


One thing to keep in mind is that:

`filter()` works on rows, and
`select()` works on columns

## Saving our results

Let's save our processed data in the `data/` directory. We'll save it as a `csv` file, which is short for *comma separated value*. This is a file type that can be easily imported into excel.


```{r}
processed_data <- smoke_complete %>%
  select(-gender) %>%
  filter(cigarettes_per_day < 20)

write_excel_csv(x = processed_data,
                path = "data/processed_data.csv")
```
