# final project
1. Find a dataset that youre interested in. It should be either in excel or csv
2. Decide on a question youre interested in
3. Explore the dataset and transform the variable
4. visualize the dataset and talk about your insights

# getting help
# https://www.biostarhandbook.com/

# Statistics
# http://www.sthda.com/english/

`tidymodels` attempts to unify all of the various modeling packages in a consistent interface.

`broom` works mostly with the output of models. One of the problems with R is that the many modeling packages are not consistent to work with. It can be just as difficult to get a p-value out of a model as it is to run it on some data! `broom` simpliflies this a lot.

# Load libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
library(readxl)
library(janitor)
library(GGally) #extends ggplot
```
# Load dataset. Assign boolean to gender (0 = females)
```{r message=FALSE}
body_comp <- read_csv('data/body_composition.csv', na="NA") %>%
  clean_names() %>%
  mutate(gender = factor(gender, levels=c("1", "0")))

head(body_comp)
```
# Pivot to give methods a column. We are losing some info here
```{r}
body_comp_long <- body_comp %>%
  pivot_longer(cols = c('dxa', 'st', 'br'),
               names_to = 'method',
               values_to = 'body_fat_percentage')
head(body_comp_long)
```
# Boxplot it
```{r warning=FALSE}
ggplot(body_comp_long) +
  aes(x = method, y = body_fat_percentage, fill = method) +
  geom_boxplot() +
  geom_jitter(color="grey")
```

# Paired T test
- Use when we are examining the mean when the subjects are the SAME
- We wouldnt use this if the subjects for each method were two separate groups
- Can only be done between two groups - we'll do dxa and st

```{r}
body_comp_dxa_st <- body_comp_long %>%
  filter(method %in% c("dxa", "st"))


tidy_output2 <- t.test(body_comp$dxa, body_comp$st, paired=TRUE) %>%
  tidy()

# same thing?
tidy_output <-
  t.test(body_fat_percentage ~ method,
        paired=TRUE,
        data=body_comp_dxa_st) %>%
    tidy()

tidy_output2
```

# Now do it for dxa and br
```{r eval=FALSE}
body_comp <- body_comp %>%
  tidyr::drop_na()

body_comp_long <- body_comp %>%
  pivot_longer(cols = c('dxa', 'st', 'br'),
               names_to = 'method',
               values_to = 'body_fat_percentage')

body_comp_dxa_sf <- body_comp_long %>%
  filter(method %in% c("dxa", "br"))

tidy_output <-
  t.test(body_fat_percentage ~ method,
       paired=TRUE,
       data=body_comp_dxa_sf) %>%
    tidy()

tidy_output
```


# Linear regression - how correlated are the three variables?

# ggpairs() - a lot of info in one plot. Keep the variables low
```{r message=FALSE}
GGally::ggpairs(body_comp,  aes(color=gender))  
```
# Linear method trendline
```{r}
body_comp %>%
  ggplot() +
  aes(x=dxa, y=st) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) #se = True will add the standard error

```


# Building linear model
```{r}
# dependent variable on left of ~
lm(dxa ~ st, data=body_comp) %>%
  tidy()
```
The predicted line through the data is:
dxa  = 0.295    +    0.903 * `st`
# Adding another variable (gender)

```{r}
copy_number <- c("1", "0", "2")


lm(dxa ~ st + gender, data=body_comp) %>%
  tidy()
```
dxa = 0.097 +   0.889 * `st`    +    0.536 * `gender0`

# Adding br
```{r}
lm(dxa ~ st + br, data=body_comp) %>%
  tidy()
```
Multicolinearity: adding the same information multiple times in the models
- Can make terms that should be significant not significant ***
- This is what happens when we add 'br' into the model (changes p value!)
