```{r prep, message=FALSE, warning=FALSE}
# load libraries
library(tidyverse)
library(readxl)
library(janitor)
library(forcats)

smoke_complete <- read_excel("data/smoke_complete.xlsx",
                             sheet=1,
                             na="NA")
```
# Advantages with using factors
1. Control ordering on axis
2. Permissible values - When do define a factor it will ignore all other levels

# 1
# Add a feature that is the a factor of the disease types
```{r}
smoke_complete2 <- smoke_complete %>%
    mutate(disease_factor =
               factor(disease,
                      levels = c("LUSC", "CESC", "BLCA", "BRCA")
                      )
           )
# Plot - notice how x-axis is in the order we defined the levels of the factor
ggplot(smoke_complete2) +
 aes(x=disease_factor, y=cigarettes_per_day) +
 geom_boxplot()
```
# 2
```{r}
character_vector <- c("LUSC", "LUSC", "BRCA", "BLCA", "BRCA", "CESC", "CESC")

factor_vector <- factor(character_vector, levels=c("LUSC", "BRCA"))

factor_vector
```
# Manipulating factors
- fct_rev() reverses the order of the levels
- fct_reorder() sorts factor by a numeric variable
- fct_collapse() collapse multiple categories into one
- fct_recode() - lets you recode values manually
- fct_other() - lets you define what categories are in "other" variable
```{r}
# fct_rev()

smoke_complete3 <- smoke_complete2 %>%
  mutate(disease_rev = fct_rev(disease_factor))
ggplot(smoke_complete3) + aes(y=disease_rev, x=cigarettes_per_day) +
  geom_boxplot()
```
```{r}
# fct_reorder()

smoke_complete3 <- smoke_complete2 %>%
  mutate(disease_reorder = fct_reorder(disease_factor, cigarettes_per_day))

ggplot(smoke_complete3) + aes(y=disease_reorder, x=cigarettes_per_day) +
  geom_jitter()
```
```{r}
# fct_collapse()

smoke_complete3 %>%
  mutate(disease_collapse = fct_collapse(
    disease_factor,
    other = c("BLCA", "CESC"),
    LUSC = c("LUSC")
  )) %>%
  tabyl(disease_collapse)
```
